---
title: "MacPlots"
author: "Gleb Furman"
---



```{r, echo = F, warning = F, message = F}
rm(list = ls())

library(dplyr)
library(ggplot2)
library(tidyr)
library(stringr)

df <- read.csv("Results/MacKinnon/20161016.csv")

df <- filter(df, !(HC == "OLS" & test %in% levels(df$test)[-5])) %>% 
  select(-X, -(B:power), -(iterations:seed), -(coef:criterion), -percent_NA) %>%
  gather(key = alpha, value = p, p.005:p.100)

df$alpha <- as.numeric(sub(".", "", df$alpha))

MCSE <- function(a, k = 20000) sqrt((a * (1-a))/k) * qnorm(.95)

HC3 <- filter(df, test == "naive", HC == "HC3") %>% 
  mutate(pHC3 = p) %>%
  select(-p, -(HC:test))

  
df <- merge(df, HC3) %>% mutate(ULa = alpha + MCSE(alpha),
                                LLa = alpha - MCSE(alpha))
```

```{r, echo = F, warning = F, message = F}
test <- levels(df$test)[!(levels(df$test) %in% "naive")]
HC <- levels(df$HC)[!(levels(df$HC) %in% "OLS")]

df %>% 
  mutate(comb = paste(HC, test)) %>%
  group_by(HC, test, n, alpha) %>%
  mutate(under = prod(p - MCSE(alpha) <= ULa),
         hc3under = prod(pHC3 <= ULa),
         underHC3 = sum(p + MCSE(alpha) >= pHC3) > 1,    #set to show if beats naive HC3 for at least one value of g
         #underHC3 = mean(p + MCSE(alpha) >= pHC3) >= .75,  #set to show if beats naive HC3 for at least 75% of g values
         beatHC3 = ifelse(hc3under == 0, under, under*underHC3)) %>%
  ungroup -> df


# Uncomment to collapse across Ns
df %>%
  group_by(HC, test, alpha) %>%
  mutate(beatHC3 = prod(beatHC3)) %>%
  ungroup -> df

```
  
The first objective was to identify which combination of HCCME and adjustment (where applicable) outperformed the naive HC3 estimator. Figures 1-4 show the performance each combination at $\alpha$ = .005, .010, .050, and .010 respectively. Lines which crossed or remained above nominal alpha plus MCSE (the dotted horizontal line) for any n were ommited from the graphs as they represented tests which were too liberal. The exception to this is the black solid line which represents the naive HC3 estimator. Figure 1 shows that at the .005 level, the HC3 overly liberal until a large to extreme amount of heteroskedasticity is introduced (g > 1). At this point the adjusted HC2 is prefered, specifically the saddle_H across all values of N, as well as KCCI_E and Satt_E which performens at the level of or better than the saddle_H at n > 20. 

```{r, echo = F, warning = F, message = F}
cmb <- expand.grid(HC[-(2:3)], test, stringsAsFactors = F)
cmb <- paste(cmb$Var1, cmb$Var2)

pv <- .005

df %>%
  filter(HC != "OLS", test == "naive") %>%
  ggplot(aes(x = g,
             y = p,
             color = HC)) +
  geom_smooth(method = "loess", se = F, size = 1) +
  geom_hline(aes(yintercept = ULa),
             linetype = "dashed") +
  geom_hline(aes(yintercept = alpha)) +
  facet_grid(alpha~n, scale = "free")

# fig0
# 
# data.frame(fig0$data %>% select(comb) %>% distinct(comb)) -> best0
# paste(best0[,1], collapse = ", ")

```
  
```{r, echo = F, warning = F, message = F}
cmb <- expand.grid(HC[-(2:3)], test, stringsAsFactors = F)
cmb <- paste(cmb$Var1, cmb$Var2)

pv <- .005

excl <- c("HC1 Rp_H", "HC2 Rp_H")

df %>%
  filter(HC != "OLS", test != "naive", beatHC3 == 1, !(comb %in% c(cmb, excl)), alpha == pv) %>%
  ggplot(aes(x = g,
             y = p,
             color = test,
             linetype = HC)) +
  geom_smooth(method = "loess", se = F, size = 1) +
  geom_hline(aes(yintercept = ULa),
             linetype = "dashed") +
  geom_hline(aes(yintercept = alpha)) +
  facet_wrap(~n)  +
  geom_smooth(data = filter(df, HC == "HC3" & test == "naive" & alpha == pv) %>% select(-HC, -test),
              color = "black",
              linetype = "solid",
              size = .2, 
              se = F) +
  coord_cartesian(ylim = c(.000, 1.25*pv)) + 
  labs(title = "Fig 1: .005 plots") + 
  scale_color_brewer(type = "qual", palette = 2) -> fig1

fig1  +
  geom_smooth(data = filter(df, HC == "HC1" & test == "RCI_H" & alpha == pv) %>% select(-HC, -test),
              color = "black",
              linetype = "solid",
              size = .2, 
              se = F) +
  geom_point()


data.frame(fig1$data %>% select(comb) %>% distinct(comb)) -> best1
paste(best1[,1], collapse = ", ")

```
  
  At $\alpha$ = .010 Figure 2 shows that HC1 with the RCI_H adjustment outperforms all other combinations at more extreme heteroskedasticity (g > 1) though the HC2 with saddle_H, saddle_E and Satt_E adjustments perform equally or better at larger ns with lower levels of heteroskedasticity. 
  
```{r, echo = F, warning = F, message = F}
cmb <- expand.grid(HC[-(2:3)], test, stringsAsFactors = F)
cmb <- paste(cmb$Var1, cmb$Var2)

pv <- .01

excl <- c("HC1 Rp_H", "HC2 Rp_H")

df %>%
  filter(HC != "OLS", test != "naive", beatHC3 == 1, !(comb %in% c(cmb, excl)), alpha == pv) %>%
  ggplot(aes(x = g,
             y = p,
             color = test,
             linetype = HC)) +
  geom_smooth(method = "loess", se = F, size = 1) +
  geom_hline(aes(yintercept = ULa),
             linetype = "dashed") +
  geom_hline(aes(yintercept = alpha)) +
  facet_wrap(~n)  +
  geom_smooth(data = filter(df, HC == "HC3" & test == "naive" & alpha == pv) %>% select(-HC, -test),
              color = "black",
              linetype = "solid",
              size = .2, 
              se = F) + 
  scale_color_brewer(type = "qual", palette = 2) +
  coord_cartesian(ylim = c(.000, 1.25*pv)) + 
  labs(title = "Fig 2: .010 plots") -> fig2

fig2

data.frame(fig2$data %>% select(comb) %>% distinct(comb)) -> best2
paste(best2[,1], collapse = ", ")

```
  Figure 3 plots show that for the first time HC3 is at or below nominal $\alpha$ = .050 when n = 20. All HCCMEs and adjustments displayed appear to perform as well as or better than HC3, with the HC1 RCI_H adjustment begining to outperform all others as g increases.
  
```{r, echo = F, warning = F, message = F}
cmb <- expand.grid(HC[-(2:3)], test, stringsAsFactors = F)
cmb <- paste(cmb$Var1, cmb$Var2)

pv <- .05

df %>%
  filter(HC != "OLS", test != "naive", beatHC3 == 1, !(comb %in% cmb), alpha == pv) %>%
  ggplot(aes(x = g,
             y = p,
             color = test,
             linetype = HC)) +
  geom_smooth(method = "loess", se = F, size = 1) +
  geom_hline(aes(yintercept = ULa),
             linetype = "dashed") +
  geom_hline(aes(yintercept = alpha)) +
  facet_wrap(~n)  +
  geom_smooth(data = filter(df, HC == "HC3" & test == "naive" & alpha == pv) %>% select(-HC, -test),
              color = "black",
              linetype = "solid",
              size = .2, 
              se = F) + 
  scale_color_brewer(type = "qual", palette = 2) +
  coord_cartesian(ylim = c(.000, 1.25*pv)) + 
  labs(title = "Fig 3: .050 plots") -> fig3

fig3

data.frame(fig3$data %>% select(comb) %>% distinct(comb)) -> best3
paste(best3[,1], collapse = ", ")

```
    Figure 4 plots show that HC3 is at or below nominal $\alpha$ = .100 for all sample sizes. At n = 20 all HCCMEs and adjustments displayed appear to perform as well as or better than HC3, though as N increases, the naive HC3 becomes more accurate at while g is low. However, as g increases, the HC1 RCI_H adjustment once again outperforms all others.
    
```{r, echo = F, warning = F, message = F}
cmb <- expand.grid(HC[-(2:3)], test, stringsAsFactors = F)
cmb <- paste(cmb$Var1, cmb$Var2)

pv <- .1

df %>%
  filter(HC != "OLS", test != "naive", beatHC3 == 1, !(comb %in% cmb), alpha == pv) %>%
  ggplot(aes(x = g,
             y = p,
             color = test,
             linetype = HC)) +
  geom_smooth(method = "loess", se = F, size = 1) +
  geom_hline(aes(yintercept = ULa),
             linetype = "dashed") +
  geom_hline(aes(yintercept = alpha)) +
  facet_wrap(~n)  +
  geom_smooth(data = filter(df, HC == "HC3" & test == "naive" & alpha == pv) %>% select(-HC, -test),
              color = "black",
              linetype = "solid",
              size = .2, 
              se = F) + 
  scale_color_brewer(type = "qual", palette = 2) +
  coord_cartesian(ylim = c(.000, 1.25*pv)) + 
  labs(title = "Fig 4: .100 plots") -> fig4

fig4

data.frame(fig4$data %>% select(comb) %>% distinct(comb)) -> best4
paste(best4[,1], collapse = ", ")

```